---
title: "Grade analysis"
author: "Leslie HUYNH"
format: html
---
https://github.com/leslie3001/r-101-grade-analysis.git 

```{r}
#| message: false
library(here)
library(vroom)
library(ggplot2)
library(dplyr)
library(tidyr)
here::i_am("grade_analysis.qmd")
```
## Question 1
```{r}
#| message: false
grades <- vroom(here("grades.csv"))
```
## Question 2
```{r}
grades |> summarise(
  min_exam=min(Exam, na.rm=TRUE),
  max_exam=max(Exam, na.rm=TRUE),
  med_exam=median(Exam, na.rm=TRUE),
  mean_exam=mean(Exam, na.rm=TRUE)
) |>
  knitr::kable()
```

## Question 3
```{r}
missing <- grades |> filter(is.na(Exam)) |> nrow()
```
`r missing` students did not take the final exam. 

## Question 4
```{r}
grades |> filter(Exam!="NA")|>
  ggplot(aes(x=Exam))+ 
  geom_bar()+
  scale_x_continuous(breaks=0:20)+
  xlab("Final exam grades")
```
This graph shows the repartition of grades at the Exam. 

## Question 5
```{r}
grades |> group_by(Group)|> summarise(number_group=n())|>
  knitr::kable()
```

## Question 6
```{r}
grades |> group_by(Group) |> 
  ggplot(aes(y=Group)) + geom_bar()
```
This graph gives the repartition of students in each group.

## Question 7
```{r}
grades |> filter(Exam!="NA") |>
  ggplot(aes(x=Group, y=Exam))+geom_boxplot()+
  ylab("Grades")
```

```{r}
grades |> filter(Exam!="NA") |>
  ggplot(aes(x=Group, y=Exam))+geom_violin()+ coord_flip()+
  ylab("Grades")
```
These two graphs shows the grades of each group. 

## Question 8
```{r}
missing_grades <- grades |> group_by(Group) |> reframe(missexam=is.na(Exam))
missing_grades |> group_by(Group) |> summarise(N=sum(missexam==TRUE)) |>
knitr::kable()
```

## Question 9
```{r}
missing_grades |> group_by(Group) |> summarise(N=sum(missexam==TRUE)) |>
ggplot(aes(y=Group, x=N)) + 
  geom_col() + 
  scale_x_continuous(breaks=0:10)+
  labs(title="Number of students who missed the Exam in each group")

```

## Question 10
```{r}
grades2 <- grades |> pivot_longer(cols=3:32, names_to="name", values_to="value")
```

## Question 11
```{r}
missing_grades2 <- grades2 |> group_by(Id) |> reframe(N=is.na(value))
missing_grades2 |> group_by(Id) |> summarise(exam_missed=sum(N==TRUE))

```
## Question 12
```{r}
missing_grades2 |> group_by(Id) |> summarise(exam_missed=sum(N==TRUE)) |> ggplot(aes(x=Id, y=exam_missed))+
  geom_col()+
  labs(title="Distribution of the number of missing grades per student")
```
```{r}
missing_grades2 |> group_by(Id) |> summarise(exam_missed=sum(N==TRUE)) |> ggplot(aes(x=exam_missed))+
  geom_bar()+
  ylab("the number of students who missed x exams")
```
Here is another representation (clearer) of the number of missing grades. Yet, one cannot know how many grades one student in particular is missing. 

## Question 13
```{r}
group_miss <- grades2 |> group_by(Group) |> filter(name=="Exam") |> reframe(N=is.na(value))
group_miss |> group_by(Group) |> summarise(N=sum(N==TRUE)) |>
knitr::kable()
```
## Question 14
```{r}
online_grades <- grades[, -c(3:13)]
online_grades_long <- online_grades |> pivot_longer(cols=3:21, names_to = "Online Exam", values_to = "grades")
```

```{r}
miss_online_grades <- online_grades_long |> group_by(Id) |> reframe(nb_miss=is.na(grades))
miss_online_grades |> group_by(Id) |> summarise(nb=sum(nb_miss==TRUE))
```
For each student, one can see how many online tests they missed. 

## Question 15
```{r}
grades15 <- grades |>
  rowwise() |>
  mutate(Missing_Online_Grades = sum(is.na(c_across(starts_with("Online_MCQ")))))
```

```{r}
grades15 |>
  ggplot(aes(x = Missing_Online_Grades, y=Exam)) +
  geom_jitter(alpha=0.5) + 
  labs(
    title = "Distribution of Exam Grades by Number conditionned by Missing Online Test Grades",
    x = "Number of Missing Online Test Grades",
    y = "Exam Grade"
  )
```
This scatter plot tries to show the correlation between the Exam grades and the number of missing online tests. 


## Question 16
```{r}
missed_table <- grades |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across(starts_with("MCQ_"))))) |>
  select(Id, Missed)
```

## Question 17
```{r}
step1 <- grades |> distinct(Id, Group)
```


```{r}
group_miss_table <- inner_join(step1, missed_table)
```
```{r}
group_miss_table |> group_by(Group) |> summarise(percent=sum(Missed==FALSE)/n()*100) |> 
  knitr::kable()
```
This table gives the proportion of students in each group that misses at least one MCQ test.

## Question 18
```{r}
percent_table <- group_miss_table |> group_by(Group) |> summarise(percent=sum(Missed==FALSE)/n()*100)
```

```{r}
mean_grades <- grades |> group_by(Group) |> mutate(mean_exam = mean(Exam, na.rm=TRUE))
```


```{r}
mean_grades <- mean_grades |> distinct(Group, mean_exam)
final_merge <- inner_join(percent_table, mean_grades)
```

```{r}
final_merge |> ggplot(aes(x=mean_exam, y=percent, color=Group)) +
  geom_point(size=5)+
  scale_x_continuous(breaks = 0:20) +
  xlab("average grade") +
  ylab("% of students in a group that misses at least 1 grade")
```
Average grade at the Exam of each group as a function of the percentage of missed exams. 
